`State and Parameter Estimation for Spatio-Temporal Neural Fields' or `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-18 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
rearranged equation 38
corrected figure6 caption change the solid and dotted line
minor edits

Alternatively, given the basis function width, the spatial cut-off frequency of the field modelled by the decomposition can be determined from spectral analysis. This is particularly useful when considering the spatially distributed basis functions, since Shannon's sampling theorem is derived for perfect point sensors. To calculate the cut-off frequency imposed by the basis function, $\boldsymbol{\nu}_{c\phi}$, we rearrange equations~\ref{eq:WidthCutOffRelationship} and \ref{eq:WidthFrequencyRelationship} giving
\begin{equation}\label{eq:CutoffFromBasisFuncWidth}
	\boldsymbol{\nu}_{c\phi}=\sqrt{\frac{\ln2}{2}}\frac{1}{\pi\sigma_{\phi}}.
\end{equation}

@@

Histograms of the parameter estimates obtained over the 100 realisations are shown in Figure~\ref{fig:Parameters}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good concordance with the actual values. 
% It should be noted that when the excitatory connectivity parameters, $\hat\theta_0$ and $\hat\theta_2$, were under (or over) estimated, the inhibitory connectivity parameter, $\hat\theta_1$, tended to be over (or under) estimated. This is demonstrated in Figure~\ref{fig:ParametersRatio} where we show the distribution of the ratios of the parameter estimates. 
\begin{table}
\centering
\begin{tabular}{cccc}
	\hline\hline Parameter & Actual Value & Mean Estimate $\pm$ Std. Dev. & Bias (\%) \\
	\hline\hline
 	$\xi$ & $0.9$ & $0.925\pm 0.004$ & $2.778$ \\
	$\theta_0$ & $10$ & $10.228\pm2.794$ & $2.28$\\
	$\theta_1$ & $-8$ & $-8.495\pm1.913$ & $-6.187$\\
	$\theta_2$ & $0.5$ & $0.545\pm0.077$ & $9$\\
	$\mid\theta_0/\theta_1\mid$ &$1.25 $ & $1.189\pm 0.0812$ &$4.88$\\
	$\theta_0/\theta_2$ &$20$ & $18.567\pm3.757$ &$7.165$ \\
\hline\hline
\end{tabular}
\caption{Actual parameter values compared to parameter estimates.}
\label{tab:Parameters estimates}
\end{table}
\begin{figure}
    \centering
    \subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
    \subfigure[][]{\label{fig:theta1dis}\includegraphics{./Graph/theta1.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameter estimates over 100
realisations. The dotted line shows the actual value of the parameters and the solid line shows the mean of the estimated parameters. \textbf{a.} The distribution of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Parameters}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). In each case, the solid and dotted lines show the locations of the true and estimated ratios respectively.}
\label{fig:ParametersRatio}
\end{figure}
The reconstructed connectivity kernels from the estimated parameters are illustrated in Figure~\ref{fig:KernelEstimates}, where the mean over the 100 realisations is shown as a dotted line and the $95\%$ confidence interval is indicated by the grey shaded region. The figure verifies that the shape of the estimated connectivity kernel is in good accordance to the true kernel. The true kernel lies within the confidence interval; however, the mean estimated kernel is more negative than the actual kernel due to the small bias on the estimates of the inhibitory parameter, $\hat\theta_1$, as seen in Figure~\ref{fig:theta1dis}.

@@

The closest competitor to the our `grey-box' integro-difference equation (IDE) based approach is perhaps the `black-box' multivariate auto-regressive (MVAR) model. Examples of MVAR models for estimating functional connectivity include Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}. A distinct advantage of the IDE-based technique proposed in this paper, is the estimates the connectivity structure are made independently of the number of sensors (assuming the field is sufficiently sampled). In addition, the number of parameters required to model the system is considerably less than a typical MVAR model. Furthermore, by using the neural field equations as a parametric form of the model, the parameters have physical meaning and therefore, can be clinically relevant. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in the longer range cortico-cortical fibres. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of a multi-scale connectivity structure in generating the characteristic rhythms of the brain. A possible method of creating more precise models with this multi-scale structure maybe to estimate anisotropic cortico-cortical connectivity using a MVAR approach and intracortical connectivity using the methods proposed in this paper.

>>>>

Alternatively, given the basis function width, the spatial cut-off frequency of the field modelled by the decomposition can be determined from spectral analysis. This is particularly useful when considering the spatially distributed basis functions, since Shannon's sampling theorem is derived for perfect point sensors. To calculate the cut-off frequency imposed by the basis function, $\boldsymbol{\nu}_{c\phi}$, we rearrange equations~\ref{eq:WidthCutOffRelationship} and \ref{eq:WidthFrequencyRelationship} giving
\begin{equation}\label{eq:CutoffFromBasisFuncWidth}
	\boldsymbol{\nu}_{c\phi}=\frac{1}{\pi\sigma_{\phi}}\sqrt{\frac{\ln2}{2}}.
\end{equation}

@@

Histograms of the parameter estimates obtained over the 100 realisations are shown in Figure~\ref{fig:Parameters}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good concordance with the actual values. 
% It should be noted that when the excitatory connectivity parameters, $\hat\theta_0$ and $\hat\theta_2$, were under (or over) estimated, the inhibitory connectivity parameter, $\hat\theta_1$, tended to be over (or under) estimated. This is demonstrated in Figure~\ref{fig:ParametersRatio} where we show the distribution of the ratios of the parameter estimates. 
\begin{table}
\centering
\begin{tabular}{cccc}
	\hline\hline Parameter & Actual Value & Mean Estimate $\pm$ Std. Dev. & Bias (\%) \\
	\hline\hline
 	$\xi$ & $0.9$ & $0.925\pm 0.004$ & $2.778$ \\
	$\theta_0$ & $10$ & $10.228\pm2.794$ & $2.28$\\
	$\theta_1$ & $-8$ & $-8.495\pm1.913$ & $-6.187$\\
	$\theta_2$ & $0.5$ & $0.545\pm0.077$ & $9$\\
	$\mid\theta_0/\theta_1\mid$ &$1.25 $ & $1.189\pm 0.0812$ &$4.88$\\
	$\theta_0/\theta_2$ &$20$ & $18.567\pm3.757$ &$7.165$ \\
\hline\hline
\end{tabular}
\caption{Actual parameter values compared to parameter estimates.}
\label{tab:Parameters estimates}
\end{table}
\begin{figure}
    \centering
    \subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
    \subfigure[][]{\label{fig:theta1dis}\includegraphics{./Graph/theta1.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameter estimates over 100
realisations. The solid line shows the actual value of the parameters and the dotted line shows the mean of the estimated parameters. \textbf{a.} The distribution of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Parameters}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). In each case, the solid and dotted lines show the locations of the true and estimated ratios respectively.}
\label{fig:ParametersRatio}
\end{figure}
The reconstructed connectivity kernels from the estimated parameters are illustrated in Figure~\ref{fig:KernelEstimates}, where the mean over the 100 realisations is shown as a dotted line and the $95\%$ confidence interval is indicated by the grey shaded region. The figure verifies that the shape of the estimated connectivity kernel is in good accordance to the true kernel. The true kernel lies within the confidence interval; however, the mean estimated kernel is more negative than the actual kernel due to the small bias on the estimates of the inhibitory parameter, $\hat\theta_1$, as seen in Figure~\ref{fig:theta1dis}.

@@

The closest competitor to our `grey-box' integro-difference equation (IDE) based approach is perhaps the `black-box' multivariate auto-regressive (MVAR) model. Examples of MVAR models for estimating functional connectivity include Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}. A distinct advantage of the IDE-based technique proposed in this paper, is the estimates the connectivity structure are made independently of the number of sensors (assuming the field is sufficiently sampled). In addition, the number of parameters required to model the system is considerably less than a typical MVAR model. Furthermore, by using the neural field equations as a parametric form of the model, the parameters have physical meaning and therefore, can be clinically relevant. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in the longer range cortico-cortical fibres. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of a multi-scale connectivity structure in generating the characteristic rhythms of the brain. A possible method of creating more precise models with this multi-scale structure maybe to estimate anisotropic cortico-cortical connectivity using a MVAR approach and intracortical connectivity using the methods proposed in this paper.
